### 2019-04-16 Thesis/thesis.tex
Nightly commit.
Finished chapter.

\section{Monads and Kleisli categories}
\label{SecMonads}

\subsection{Monads}

@@

  Going the other way, if $\M$ is monoidal closed with inner hom $\implies$, and if $z$ is a comonoid in $\M$, then the \emph{reader monad} $R_z$ is given by $R_z y = z \implies y$.  
  Then the monadic coherences
  \begin{mathpar}
    m_y \from z \implies z \implies y \to z \implies y
    \and
    u_y \from y \to z \implies y
  \end{mathpar}
  are induced from the comonoid structure on $z$.
  This second example is particularly important in Cartesian closed categories, in which every object has the structure of a comonoid.
\end{example}
\begin{example}
  If $\adjunction{\C}{L}{R}{\D}$ is an adjunction with counit $\epsilon\from LR\to 1$ and unit $\eta\from 1 \to RL$, then the composite $RL\from \C \to \C$ has the structure of a monoid on $\C$, where the multiplication and unit are given by
  \begin{mathpar}
    R\epsilon L \from RLRL \to RL
    \and
    \eta \from 1 \to RL\,.
  \end{mathpar}
  We will see in the next section that every monad is induced by an adjunction in this way.

@@

\subsection{Kleisli Categories}

@@

\subsection{Denotational Semantics}

From now till the end of the chapter, we fix an (order-enriched) Cartesian closed category $\G$ that admits a denotational semantics of Idealized Algol satisfying Computational Adequacy and in which every compact element is definable.
The prototypical example, of course, will be the category of games and visible strategies, but we will not exploit any properties of this model beyond the ones we have already mentioned, mentioning it only in examples where appropriate.

@@

\subsection{Operational Semantics}

@@

\subsection{Soundness}
\label{SecSoundness}

@@

\begin{definition}
  Let $n$ be a natural number.  
  We define terms $\test_n\from \nat \to \com$ inductively by
  \begin{mathpar}
    \test_0 = \lambda m.\IfO m \Then \skipp \Else \Omega
    \and
    \test_{n+1} = \lambda m.\IfO m \Then \Omega \Else \test_n (\pred m)\,.
  \end{mathpar}
  So $\test_n$ converges if its input evaluates to $n$ and diverges otherwise.

@@

  We can prove this using the following derivation tree.
  \tiny
  \[
    \inferrule*
    {
      \inferrule*[sep=1.7em]
      {
        \inferrule*[left={Lem. \ref{LemSoundness}}]
        {
        }
        {
          v,(v\mapsto 0)\ts M[v\gets\suc!v;\tr_{u^\top} !v/\ask_x]\converges \skipp,(v\mapsto |u|)
        }
        \\
        \inferrule*
        {
          \inferrule*
          {
          }
          {
            v,(v\mapsto |u|) \ts v \converges v,(v\mapsto |u|)
          }
        }
        {
          v,(v\mapsto |u|)\ts !v \converges |u|,(v\mapsto |u|)
        }
      }
      {
        v,(v\mapsto 0) \ts M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X];v \converges |u|,(v\mapsto |u|)
      }
    }
    {
      ,()\ts\new (\lambda v.M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X];!v) \converges |u|,()
    }\qedhere
    \]
  \normalsize
\end{proof}

@@

\subsection{Computational Adequacy}

@@

\subsection{Full abstraction}

@@

\subsection{Comparison with Ghica's slot games}

@@

\subsection{Alternative reduction rules - may testing}

@@

\subsection{Alternative reduction rules - must testing}

@@

\subsection{Full Abstraction for Nondeterminism and Must Testing}

@@

\subsection{The intrinsic equivalence relation in $\G_X$}

We conclude this section by examining the intrinsic equivalence relation in $\G_X$.

>>>>

\tableofcontents
\newpage

\chapter{Monads and Kleisli categories}
\label{ChapMonads}

\section{Monads}

@@

  Going the other way, if $\M$ is monoidal closed with inner hom $\implies$, and if $z$ is a comonoid in $\M$, then the \emph{reader monad} $R_z$ is given by $R_z y = z \implies y$.  
  Then the monadic coherences
  \begin{mathpar}
    m_y \from z \implies z \implies y \to z \implies y
    \and
    u_y \from y \to z \implies y
  \end{mathpar}
  are induced from the comonoid structure on $z$.
  This second example is particularly important in Cartesian closed categories, in which every object has the structure of a comonoid.
\end{example}
\begin{example}
  If $\adjunction{\C}{L}{R}{\D}$ is an adjunction with counit $\epsilon\from LR\to 1$ and unit $\eta\from 1 \to RL$, then the composite $RL\from \C \to \C$ has the structure of a monoid on $\C$, where the multiplication and unit are given by
  \begin{mathpar}
    R\epsilon L \from RLRL \to RL
    \and
    \eta \from 1 \to RL\,.
  \end{mathpar}
  We will see in the next chapter that every monad is induced by an adjunction in this way.

@@

\section{Kleisli Categories}

@@

\section{Denotational Semantics}

Having given an overview of the general theory of monads and Kleisli categories, we now examine the relationship between Kleisli categories and Full Abstraction.  From now till the end of the chapter, we fix an base Cartesian closed category $\G$ that admits a denotational semantics of Idealized Algol satisfying Computational Adequacy.  
In addition, we require that $\G$ can be interpreted as being enriched in directed-complete partial orders, in such a way that any compact morphism between the denotation of IA types is the denotation of some IA term.
The prototypical example, of course, will be the category of games and visible strategies, but we will not exploit any properties of this model beyond the ones we have already mentioned, mentioning it only in examples where appropriate.

In addition, we will specialize to reader monads.  
The rationale behind this is twofold: firstly, we do not wish to assume too much about the underlying category $\G$ beyond the fact that it gives us a good model of Idealized Algol.  
Secondly, in order to get a compositional semantics for the $\lambda$-calculus, it is important that the Kleisli categories we consider are themselves Cartesian closed.

@@

\section{Operational Semantics}

@@

\section{Soundness}
\label{SecSoundness}

@@

\begin{definition}
  Let $n$ be a natural number.  
  We inductively define terms $\test_n\from \nat \to \com$ by
  \begin{mathpar}
    \test_0 = \lambda m.\IfO m \Then \skipp \Else \Omega
    \and
    \test_{n+1} = \lambda m.\IfO m \Then \Omega \Else \test_n (\pred m)\,.
  \end{mathpar}
  So $\test_n$ converges if its input evaluates to $n$ and diverges otherwise.

@@

  We can prove this using the following derivation tree.
  \[
    \inferrule*
    {
      \inferrule*[sep=1.7em]
      {
        \inferrule*[left={Lem. \ref{LemSoundness}}]
        {
        }
        {
          v,(v\mapsto 0)\ts M[v\gets\suc!v;\tr_{u^\top} !v/\ask_x]\converges \skipp,(v\mapsto |u|)
        }
        \\
        \inferrule*
        {
          \inferrule*
          {
          }
          {
            v,(v\mapsto |u|) \ts v \converges v,(v\mapsto |u|)
          }
        }
        {
          v,(v\mapsto |u|)\ts !v \converges |u|,(v\mapsto |u|)
        }
      }
      {
        v,(v\mapsto 0) \ts M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X];v \converges |u|,(v\mapsto |u|)
      }
    }
    {
      ,()\ts\new (\lambda v.M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X];!v) \converges |u|,()
    }\qedhere
    \]
\end{proof}

@@

\section{Computational Adequacy}

@@

\section{Full Abstraction}

@@

\section{Comparison with Ghica's slot games}

@@

\section{Alternative Reduction Rules - May Testing}

@@

\section{Alternative Reduction Rules - Must Testing}

@@

\section{Full Abstraction for Nondeterminism Under Must Testing}

@@

\section{The Intrinsic Equivalence Relation}

We conclude this chapter by examining the intrinsic equivalence relation in $\G_X$.
